\section{Evaluation}
%ours, RPL + ContikiMAC, MiCMAC
The results of our multichannel RPL protocol is compared against single-channel tree protocol and MiCMAC.

\subsection{Experimental Setup}
%//methodology? key metrics?
We evaluate the protocol in Cooja simulated environment with emulation of TMote sky nodes that feature the CC2420 transceiver, a 802.15.4 radio. The nodes run on IPv6, using UDP with standard RPL and 6LoWPAN protocols. The network consists of 16 nodes are used to run the simulation where we have 1 border router node, 1 interference node, and 14 duty cycled nodes that act as UDP clients to send packets to LPBR. We simulated a controlled interference node that generates semi-periodic bursty interference to resemble a simplified WiFi or Bluetooth transmitter on channel 26; which is the channel LPBR is listening on. The interference model that we use is described in \cite{Boano:2010:MSM:2127940.2127963}. 

//!!!why 1 interference only?
We had an interference on one channel to test our protocol in choosing a good channel to change into. The simulator, Cooja has a memory leak problem which unable us to use more interference on different channels as we would run out of memory to run the simulation for 60 minutes. We plan to run our protocol on the testbed as our future work to test on several interference channels.

RPL border router is used in order to move most processing decisions on a PC as it has more RAM and better processing capabilities than a sensor. TelosB has limited RAM and ROM of (ram value?), (rom value). By using a border router, this allows channel changing to be decided in real time. 

%RPL border router is used to interface a regular IPv6 with a regular RPL.

%(explain border router etc????) that always stays on and it is the sink of the tree, also the root node of RPL. There are 14 duty cycled nodes that acts as UDP clients that send packets to LPBR. We simulated a controlled interference node that generates semi-periodic bursty interference on channel 26; which is the channel LPBR is listening on. 

%several channels, mainly on the parents channel as the packets travel upward to measure the benefits of multi-channel operation of changing channel when failure happens within a given period. 

%- throughput?

//!!! only need 1 metric; to be changed!
We evaluate multichannel RPL variant using two performance metrics: end to end packet delivery and latency. In end to end packet delivery, the transmission success rate is calculated from the sender to the receiver over multiple hops. The latency, time difference from sending to receiving is also calculated based on Cooja log time. %We compare our multichannel protocol with Contiki's default ContikiMAC and RPL, and MiCMAC.

We run the simulation for a duration of 60 minutes to send 700 packets; 50 packets for each node. RPL runs the initial network setup for a few minutes before it is stable. We set the RPL setup time to be 5 minutes before our multichannel protocol runs for 10 minutes. After 15 minutes, the client nodes will send a normal packet periodically every 30-60 seconds to LPBR. This is done in order to avoid collision of the nodes sending at the same time. The simulation is repeated 3 times.

%We evaluate multichannel RPL variant using three performance metrics: end to end packet delivery, latency and duty cycle. In end to end packet delivery, the transmission success rate is calculated from the sender to the receiver over multiple hops. The latency, time difference from sending to receiving is also calculated based on Cooja log time. Contiki's energy profiler is used to measure the duty cycle where the radio usage time in the total run is calculated.



\subsection{Effect of Multi-channel}
//with existings - better? worse? what about RAM, ROM used?

We vary the interference rate, which is referred as \textit{clear\_time} in \cite{Boano:2010:MSM:2127940.2127963} to 0\% (no interference), 25\% (extreme), 50\% (moderate) and 75\% (mild). The test is done to evaluate our protocol behaviour in different interference rate and to compare the result with a single channel case. Figure (xxxxx) shows the averaged results from three runs that we did. We observed that during high interference and moderate interference, when the LPBR generate a random two hop channel for a node to change into, the receiving node will probe on the channel. It will either timed out or the probing messages received are less than a threshold that allows for the node to change it's listening channel to the new channel. This is as expected as our protocol checks the channel each time before deciding on the new channel to avoid interference channel. By doing this, we can be sure that the node listening channel is a good channel. This enable us to use all available channels without blacklisting any channel until we are sure it is a bad channel through our probing process. The channel quality table is built at the LPBR that over time, can be used to learn good and bad channels based on several probing processes. 

In the mild interference case, all probing messages are received even though there is interference in that channel. This is because, the probing gives good result which means that the channel can be used. As the interference rate is mild, all packets are received. This is also the case with a single channel. The interference does not affect the transmissions as the tests were done using MiCMAC and CSMA. The nodes will check the channel, backoff and retransmit as necessary.

In the single channel, the node does not have enough time to recover from the interference and drops all packets. Figure (xxxx)shows that there are more packets drop over time and it stops receiving packets as it doesn't have enough buffer to store the incoming packet and the channel becomes congested. However, as the interference rate increases, the single channel performance improves as it has more time to recover.



\subsection{Resilience to External Interference}